Biotic invasions reduce the manageability of mutualistic networks

### 2016-06-12 paper/legends.Rmd
add figure legends

\documentclass[a4paper,10pt]{article}
\usepackage[margin=22.5mm]{geometry}

% abstract + title + authors 
\usepackage{abstract}
\renewcommand{\abstractnamefont}{\normalfont\bfseries}
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} 
\usepackage{titlesec}
\usepackage{authblk}
\usepackage{datetime}
\newdateformat{usvardate}{
\monthname[\THEMONTH] \ordinal{DAY}, \THEYEAR}

% keywords
\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}

% math
\usepackage{amssymb}
\usepackage{newtxmath}
\usepackage{xfrac}

% manuscript looking document
\usepackage{setspace}
\usepackage{lineno,xcolor}
\setlength{\parskip}{0.5em}

% comments
\usepackage[draft, textsize = scriptsize, color = gray]{todonotes}

% graphics
\usepackage{graphicx}
\usepackage{float}
\usepackage{sidecap}
\sidecaptionvpos{figure}{t}
\usepackage[font=small]{caption}
\usepackage{subfig}
\usepackage{caption}

% table
\usepackage{booktabs}

% box
\usepackage{tcolorbox}

% references & links
\usepackage[
backend=bibtex,
style= authoryear,
citestyle = authoryear,
url=false,
doi=true,
isbn=false
]{biblatex}
\addbibresource{references.bib}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    citecolor=magenta,
    linkcolor=magenta
}
\usepackage{doi}
\begin{document}
\renewcommand*{\bibfont}{\small}
% line numbers
\linenumbers
\setlength\linenumbersep{15pt}
\renewcommand\linenumberfont{\normalfont\footnotesize\sffamily\color{gray}}

\doublespacing
%\onehalfspacing

% title
% Article title
  \title{
  {\LARGE \textsc{Biotic invasions reduce the manageability of mutualistic networks}}}

% Authors and Affiliations
  \author{\large E. Fernando Cagua}
  \author{\large Kate L. Wootton}
  \author{\large Daniel B. Stouffer}
  \affil{\normalsize Center of Integrative Ecology, School of Biological Sciences, University of Canterbury}
 \date{}

\maketitle

Alternative title:

\textsc{Network manageability in the context of species invasion}

<<libraries, echo = FALSE, warning = F, message = F>>=
library(magrittr)
library(knitr)
library(ggplot2)
library(glmulti)
library(Hmisc)
library(lme4)
opts_knit$set (root.dir=normalizePath('../'))
@

<<auxiliary functions, echo = FALSE, warning = F, message = F>>=
"./code/V2.0/functions" %>% 
	list.files(full.names = T) %>%
	plyr::l_ply(source)
@

<<plot theme, echo = FALSE, warning = F, message = F>>=
fer_theme <- theme_bw() +
	theme(text = element_text(family = "Times"),
	      title = element_text(size = 7.5, hjust = 0),
	      legend.title = element_text(size = 9),
	      legend.text = element_text(size = 8),
				axis.text = element_text(size = 8),
				axis.title = element_text(size = 9, hjust = 0.5), 
				strip.text = element_text(size = 9, hjust = 0), 
				strip.background = element_blank(),
				plot.margin = grid::unit(c(5, 0, 2, 0), "mm"))
@

<<read data and results, echo = FALSE, warning = F, message = F>>=
net <- "./data/V2.0/networks" %>%
	read_networks()
meta <- readr::read_csv("data/ntw_info.csv") %>% dplyr::tbl_df() %>%
	dplyr::rename(count = method)
n_matched <- readRDS("./data/V2.0/n_matched.rds")
a_m <- readRDS("./data/V2.0/n_matched_all_types.rds") 
r_w <- readRDS("./data/V2.0/random_weights.rds") %>% dplyr::tbl_df() %>%
  dplyr::rename_("net_name" = ".id")
r_d <- readRDS("./data/V2.0/random_directions.rds") %>% dplyr::tbl_df() %>%
  dplyr::rename_("net_name" = ".id")
s_p <- readRDS("./data/V2.0/species_properties.rds") %>% dplyr::tbl_df()
# mod <- readRDS("./data/V2.0/species_models.rds")
driver <- c("./data/V2.0/matching_frequency") %>% 
  list.files(full.names = T) %>%
  lapply(list.files, full.names = T) %>% 
  unlist() %>% 
  lapply(read_allRDS)
prop <- expand.grid(c("asymmetry", "max_dep"), 
                    c("scaled_FALSE", "scaled_TRUE")) %>% plyr::alply(1)
driver <- mapply(function(x, y){
  attr(x, "type") <- as.character(y[,1])
  attr(x, "scaled") <- stringr::str_split(as.character(y[,2]), "_") %>% unlist() %>% extract(2) %>% as.logical()
  return(x)
}, driver, prop)
driver_bi <- "./data/V2.0/matching_frequency_bi" %>% 
  read_allRDS()
mod <- readRDS("./data/V2.0/detailed_species_models.rds")
mo <- readRDS("./data/V2.0/detailed_species_models_data.rds")
@

\section*{Introduction}

Complex systems, such as ecological communities, are characterised by non-linear relationships in which the whole is greater than the sum of its parts \todo{add refs}. This complex systems approach has led to the development of a rich suite of analytical and simulation tools with which to understand the role of species when embedded in an intricate network of interactions\todo{add refs}. However, in contrast to other types of complex systems like neural, intra-organizational, or trust neworks, ecological networks do not have a hierarchical or distributed structure but rather tend to be closed systems in which resources recirculate\todo{add ref}. As such, ecosystems contain multiple feedbacks that allow them to self-regulate and, therefore, function relatively independent from external stimulation\todo{add ref}. This complex propertes have hindered, or at least complicated, our ability to find management solutions to the many problems ecological communities face. A framework that allow us to explain, predict, and manage ecological communities, particularly when they are faced with perturbations, needs to be able to account for their complex structure, and the dynamics that determine the state of each species---and the community as a whole. Control theory might be a good candidate framework as it is designed to deal with determining and supervising the behavior of a dynamical system and is well equiped to deal with the many feedbacks presents in ecological communities\todo{add ref}.

Research stemming from control theory has established a strong link between the structure of complex networks and their controlability, and that, in principle, it is possible to alter a ecological community's composition by modifying the abundances of only a few species \autocite{Liu2011, Cornelius2013}. Based on this work, it has been suggested that the degree distribution of ecological networks makes them inherently difficult to control \autocite{Liu2011, Ruths2014}. This theoretical result is in agreement with many examples in which the management and restoration of ecological communities is more challenging than aticipated. For instance, biotic invasions, as a major drivers of global change, can induce dramatic changes on the patterns of interactions that determine the structure of ecological networks \autocite{Baxter2004, Tylianakis2008, Ehrenfeld2010a}. The difficulties of returning invaded communities to a predisturbance state can be evidenced by the limited amount of success of restoration projects \autocite{Suding2004, Rodewald2015, Smith2016}. Although it is a double challenge, understanding how differences in network structure before and after the invasion impact our ability to manage them is a first step towards an informed recovery.

A control theoretical approach not only might provide insight to conservation but also for ecological science: by explicitly taking into account the extent to which changes in the abbundances of one species may ripple trough the community, applying control theory to ecological networks can provide an indication of which species are important from a dynamic perspective, and therefore crucial when attempting to alter (or maintain) the ecosystem state. Nevertheless, although it has been prevoiously used to highlight the disproportionate influence that humans have on marine food-webs \autocite{Isbell2013}, control theory has has rarely been aplied in an ecological context. One reason might be that the control of complex networks is still a young field and the tools necessary to account for variation on the strength of interspecific effects, which are characteristic of ecological networks, are not there.

Here, we expand previous theory of the control of complex systems and use a set of ten paired invaded and uninvaded plant-pollinator communities to investigate the link between network structure and our ability to manage them. We focus on biotic invasions because they can produce tractable changes in the structure of ecological networks. These changes can be particularly pronounced in mutualistic networks of plants and pollinators where biotic invasions have been shown to modify the strength of species interactions and the degree of network nestedness and connectivity \autocite{Olesen2002, Aizen2008, Bartomeus2008, Vila2009, Traveset2013}. In addition plant-pollinator networks provide an ideal framework to answer these questions. On one hand, community networks that quantify relative levels of interaction are readily available. On the other, the bipartite nature of pollination networks makes it is possible to simplify assumptions of how these interactions translate into interspecific effects. Specifically we ask two key questions framed in the context of a plant invasion. First, we quantify the manageability of pollination communities. Second, we ask whether some species are more important than others at driving the population dynamics of the community and which factors determine this importance.

\section*{Theory of community manageability}

\subsection*{Maximum matching}

The number of nodes necessary to fully control a complex network can be calculated by counting the number of unmatched nodes in the network's maximum matching \autocite{Liu2011}. In a directed network, a given matching is a subset of links in which no two links share a common starting node or a common ending node; a node is then matched if it is the ending node of one of the links in the matching and unmatched otherwise(\autoref{fig:explain_mm}a). A given matching is then maximum if both \textit{(i)} the number of matched links---known as matching size---is maximal and \textit{(ii)} the sum of the weights of the matched links---known as matching weight---is the largest possible among all possible matchings of that size \autocite{West2001}. To find the maximum matching, we it is useful to adopt an alternative bipartite representation of the network in which the two levels indicate the outgoing and incoming links to each node (\autoref{fig:explain_mm}b). Finding a matching in this alternative representation is equivalent to finding a set of links such that each node on the top level is matched to at most one node on the bottom level, and vice versa \autocite{West2001, Csardi2006a}.

\begin{figure}
\centering
<<fig explain mm, echo = FALSE, warning = F, message = F, results = "asis", fig.height=1.6, fig.width=6.2>>=

sp_names <- c("a[1]", "a[2]", "a[3]", "p[1]", "p[2]")

adj_matrix <- c(0,0,0,1,0,
								0,0,0,1,1,
								0,0,0,0,0,
								0,0,1,0,0,
								0,0,1,0,0) %>%
	matrix(nrow = 5, byrow = T) %>%
	`rownames<-`(sp_names) %>%
	`colnames<-`(sp_names)

net <- adj_matrix %>% 
	igraph::graph_from_adjacency_matrix(mode = "directed") 
igraph::V(net)$type <- c(F, F, F, T, T)
igraph::E(net)$weight <- 1

layoutCoordinates <- cbind(c(0,1,2,0.5,1.5), 
													 c(1,1,1,0,0)) %>%
	`colnames<-`(c("x", "y"))

adj_list <- adj_matrix %>% 
	reshape2::melt() %>% 
	dplyr::filter(value > 0)

# Generate an edge path for each pair of connected nodes
n <- 17
allEdges <- lapply(1:nrow(adj_list), edgeMaker, len = n, curved = FALSE)
allEdges <- do.call(rbind, allEdges) %>%
	dplyr::filter(Sequence != n) %>%
	dplyr::mutate(Group = factor(Group),
				 type = Group)

new_lev <- c("m", "u", "m", "m", "u")
levels(allEdges$type) <- new_lev

nodes <- data.frame(layoutCoordinates) %>%
	dplyr::mutate(spe = sp_names,
								spe_c = c("m", "m", "u", "m", "u"), 
								pos = rep(c(0.3, -0.3), c(3,2)))

p1 <- ggplot(allEdges) +  # Pretty simple plot code
	geom_path(aes(x = x, y = y, group = Group, colour = type, size = type), 
						arrow = arrow(type = "closed", angle = 15, length = grid::unit(0.1, "inches"))) + # and taper
	geom_point(data = nodes,  # Add nodes
						 aes(x = x, y = y , fill = spe_c), size = 5, pch = 21,
						 colour = "black") + 
	geom_text(data = nodes, aes(x = x, y = y + pos, label = spe), size = 3, parse = T, family = "Times") +
	scale_fill_manual(values = c("#636363", "white")) + 
	scale_colour_manual(values = c("#636363", "#bdbdbd")) +
	my_theme + xlab("") + ylab("") + 
  	xlim(c(-0.09, 2.09)) + ylim(c(-0.48, 1.48)) +
 theme(plot.margin=unit(c(4,4,-4,-4),"mm")) +
	scale_size_manual(values = c(0.5, 0.5)) +
  ggtitle("a) directed network")

net <- digraph_bipartite(net)
adj_matrix <- net %>% igraph::as_adjacency_matrix(sparse = F, type = "upper")
layoutCoordinates <- cbind(rep(0:4, 2), rep(c(1,0), each = 5)) %>%
	`colnames<-`(c("x", "y"))

adj_list <- adj_matrix %>% 
	reshape2::melt() %>% 
	dplyr::filter(value > 0)

# Generate an edge path for each pair of connected nodes
n <- 100
allEdges <- lapply(1:nrow(adj_list), edgeMaker, len = n, curved = FALSE)
allEdges <- do.call(rbind, allEdges) %>%
	dplyr::filter(Sequence != n,
								Sequence != 1) %>%
	dplyr::mutate(Group = factor(Group),
				 type = Group)

new_lev <- c("m", "m", "u", "m", "u")
levels(allEdges$type) <- new_lev

nodes <- data.frame(layoutCoordinates) %>%
	dplyr::mutate(spe = rep(sp_names, 2),
								spe_l = c(paste0(sp_names, " %->% phantom(0)"), paste0(sp_names, " %<-% phantom(0)")),
								spe_c = c(c("m", "m", "u", "m", "u"), rep("u",5)), 
								pos = rep(c(0.3, -0.3), each = 5))

p2 <- ggplot(allEdges) +  # Pretty simple plot code
	geom_path(aes(x = x, y = y, group = Group, colour = type, size = type)) + # and taper
	geom_point(data = nodes,  # Add nodes
						 aes(x = x, y = y, fill = spe_c), size = 5, pch = 21,
						 colour = "black") + 
	geom_text(data = nodes, aes(x = x + 0.08, y = y + pos, label = spe_l), parse = T, size = 3, family = "Times") +
	# coord_flip()  + scale_x_reverse() + scale_y_reverse() +
	scale_fill_manual(values = c("#636363", "white")) + 
	scale_colour_manual(values = c("#636363", "#bdbdbd")) +
	xlim(c(-0.12, 4.12)) + ylim(c(-0.48, 1.48)) +
	my_theme + xlab("") + ylab("") +
	scale_size_manual(values = c(0.5, 0.5)) +
  ggtitle("b) alternative bipartite representation")

multiplot(p1, p2, layout = matrix(nrow = 1, c(1,1,1,2,2,2,2,2)))
@
\caption{\label{fig:explain_mm} Finding a maximum matching in a complex network. a) We start with a directed network  in which the direction of the link represent the direction of control. This is, a link from $a_1$ to $p_1$ indicates that the state of $p_1$ (density/abundances in an ecological context) are influenced by the state of $a_1$. b) To facilitate the computation of the maximum matching, we can use an alternative representation of the directed network in which each node in (a) is represented by two nodes that indicate their outgoing and incoming links. In both (a) and (b) we show one of the possible maximum matchings of a network (dark grey links) that defines a set of matched ($p_1$, $p_2$, and $a_3$), unmatched ($a_1$ and $a_2$) and superior nodes ($a_1$, $a_2$, and $p_1$; dark grey circles). The minimum number of species neccessary to control this network equates to the number of unmatched nodes $n_d = 2$, which gives a manageability of $m = 0.8$.}
\end{figure}

Given a maximum matching, the link to network controllability has an intuitive basis. In control theory, a network is fully controlable when each node has a ``superior'', that is, an input that drives its dynamics. Consequently, each matched nodes can be controled by the node pointing to it. Unmatched nodes, on the other hand, have no superior in the network and must be directly controlled by external inputs \autocite{Liu2011}. In an ecological context, this is what we refer to as management interventions. Building on this, systems that require a large number of external inputs are more difficult or costly to control. Therefore, the manageability $m$ of a community can be approximated by the proportion of species on which we would need to apply management interventions in order to gain full control of the abundances of \textit{all} species in the ecological community $$m = 1- \frac{n_d}{s},$$ where $m$ ranges between zero and one, $n_d$ is the number of species needed to gain full control and $s$ is the total number of species in the community.

\subsection*{Relative importance}

The manageability of a community $m$ provides an indication of the magnitude of $n_d$ but not the identity of the species that compose this set. This is rather important since maximum matchings are often not unique in for a given network and different sets of species of size $n_d$ could potentially be used to control the network \autocite{Liu2011}. Species vary in terms of the frequency with which they are classified as superior nodes (\autoref{fig:explain_mm}), and this frequency provides an indication of their relative importance in driving the state of the community. Ecologically, these distinctions are crucial because resources are limited and therefore ecological inteventions should be focused on species that might provide the largest impact. As superior nodes \autoref{fig:explain_mm} drive the dynamics of other species and are therefore important for the controlability of the network because they either require external input or because the are part of the chain that transmit the external imputs to other species, we therefore call \textit{driver species} those that are identified as being a superior node in all possible matchings, and hence those most likely to have a disproportionate effect on the abundances of other species in their community.

As previously described, a maximum matching in a weighted network is one whose links have the largest possible total weight among all possible matchings. One way to calculate the species importance, and identify the driver species, would be to consider only maximum matchings. However empirical intraction strengths are to some extent stochastic and depend on proximate factors such as sampling method and intensity \autocite{Rezende2007}. This means that in weighted networks there is a number of matchings that have maximal size, but are not maximum matchings because their total weight is not maximal. We therefore relaxed the matching weight requirement and computed all possible matchings in the network that have the same maximal size while allowing for some tolerance regarding the weight. To compute all the matchings with maximal size, we first generated the network's line graph---another representation in which nodes are the links of the original network and links represent common ending points \autocite{West2001}. We then calculated the complement of the line graph: a graph in which nodes are connected if and only if they are not connected in the line graph. All possible matchings of maximal size can be found by enumerating all cliques (subgraphs in which every node is linked to every other) of maximum size, which equates to $1-n_d$m in this complement graph \autocite{Csardi2006a, Gutin2013}.

\section*{Methods}

In the last section, we outlined the theoretical and computational underpinings of our approach. Now we illustrate its empirical application using ten pairs of plant-pollinator communities for which we explored their manageability and the relative importance of their constituent species. The networks were constructed from previously published visitation data collected from pollination communities in Bristol, UK \autocite{Lopezaraiza-Mikel2007} and in Cap de Creus National Park, Spain \autocite{Bartomeus2008}. Notably, each network pair was composed of a community invaded by a plant and a community ``free'' of the invasive species (Table S1). In the four British uninvaded communities, the non-invaded plots were obtained by experimentally removing all the flowers of the invasive species \textit{Impatients grandulifera}. In contrast, the Spanish uninvaded communities were obtained from plots that had not yet been colonised by either of the invasive species \textit{Carpobrotus affine acinaciformis} or \textit{Opuntia stricta}. The structure of these networks was determined using visitation frequency which has been shown to be an appropriate surrogate for interspecific effects in pollination networks \autocite{Vazquez2005, Bascompte2006}. To examine whether this decision would influence our results, we also evaluated the effect of using pollinator efficiency or importance as alternative measures of species interactions in a different dataset \autocite{NeEman2010, Ballantyne2015}, and we found quantitatively results for all of these options (see Supplementary Information).

\subsection*{Community manageability}

Given the different networks, we started by quantifying their manageability. To do so, we calculated the maximum matching of the corresponding pollination network, and estimated the minimum number of species that need directed intervecntions to fully control the species abbundances in the community. If a network had more than one component (two species are in different components if there exists no path between them), we only analysed the largest one.

\subsubsection*{Weighting \& directing links}

Recall that the aforementioned maximum matching algorithm requires a directed network in which a link between species $i$ and $j$ pointing to species $j$ indicates that the abundance of $j$ can be affected by the abundance of $i$. This implies that we need first to identify a directionality for the links between species that in consistent with the dynamics of the comunity. In some ecological networks, establishing a directionality is relatively straightforward, for example when links represent biomass transfer or energy flow. Interspecific effects in pollination networks, however, are not strictly directed since the benefit is mutual between interacting species. Nevertheless, the relative extent to which a given pair of interacting species affect each other can be quantified by the magnitude of the mutual dependence \autocite{Bascompte2006}. The dependency of plant $i$ on pollinator $j$, $d_{ij}$, is the proportion of the visits coming from pollinator $j$ compared to all pollinator visits to plant $i$. Likewise, the dependency of pollinator $j$ on plant $i$, $d_{ji}$, is the ratio of the visits by pollinator $j$ to plant $i$ and all visits of pollinator $j$. As the dependencies are bidirectional, adopting this assumption generates a weighted bipartite network in which interacting pairs are connected by two directed links (\autoref{fig:explain_cd}b).

Mutual dependencies have been shown to be highly assymetric in natural communities \autocite{Bascompte2006}; in other words, if a plant species is largely dependent on a pollinator species, then that pollinator tends to depend rather weakly on the plant. This suggests that it may be possible to relax the assumption of bidirectional dependences and simplify the network further such that interacting species are only connected by one directed link when mutual dependencies are assymetric (\autoref{fig:explain_cd}c). The extent to which a species $i$ affects species $j$ relative to the extent to which species $j$ affects species $i$ can be summarised by the interaction asymmetry given by $$a(i,j) = \frac{\left | d_{ij}-d_{ji} \right |}{\textup{max}\left ( d_{ij}, d_{ji} \right )}.$$

Employing an unidrectional interaction asymmetry, as a surrogate of bidirectional mutual dependences, dramatically reduces the number of links in the network, which provides substantial reduction of the time and memory required for the computation of of all posible maximum matchings in a network, particularly for large networks (see ``relative importance of species'' subsection below).

\begin{figure}
\centering
<<fig explain control direction, echo = FALSE, warning = F, message = F, results = "asis", fig.height=1.5, fig.width=6.2>>=
sp_names <- c("a[1]", "a[2]", "a[3]", "p[1]", "p[2]")

adj_matrix <- c(0,0,0,0,0,
								0,0,0,0,1,
								0,0,0,0,0,
								1,1,1,0,0, 
								0,0,1,0,0) %>%
	matrix(nrow = 5, byrow = T) %>%
	`rownames<-`(sp_names) %>%
	`colnames<-`(sp_names)

net <- adj_matrix %>% 
	igraph::graph_from_adjacency_matrix(mode = "directed") 
igraph::V(net)$type <- c(F, F, F, T, T)
igraph::E(net)$weight <- 1

layoutCoordinates <- cbind(c(0,1,2,0.5,1.5), 
													 c(1,1,1,0,0)) %>%
	`colnames<-`(c("x", "y"))

adj_list <- adj_matrix %>% 
	reshape2::melt() %>% 
	dplyr::filter(value > 0)

# Generate an edge path for each pair of connected nodes
n <- 17
allEdges <- lapply(1:nrow(adj_list), edgeMaker, len = n, curved = FALSE)
allEdges <- do.call(rbind, allEdges) %>%
	dplyr::filter(Sequence != n) %>%
	dplyr::mutate(Group = factor(Group),
				 type = Group)

new_lev <- c("m", "u", "m", "m", "u")
levels(allEdges$type) <- new_lev

nodes <- data.frame(layoutCoordinates) %>%
	dplyr::mutate(spe = sp_names,
								spe_c = c("m", "m", "u", "m", "u"), 
								pos = rep(c(0.3, -0.3), c(3,2)))

weights <- allEdges %>%
  dplyr::filter(Sequence == 12) %>%
  dplyr::mutate(weight = c(1, 2, 1, 1, 2))

p1 <- ggplot(allEdges) +  # Pretty simple plot code
	geom_path(aes(x = x, y = y, group = Group, size = type), colour = "#bdbdbd") + # and taper
	geom_point(data = nodes,  # Add nodes
						 aes(x = x, y = y ), size = 5, pch = 21,
						 colour = "black", fill = "white") + 
	# geom_text(data = nodes, aes(x = x, y = y + pos, label = spe), size = 3, parse = T, family = "Times") +
  geom_point(data = weights, aes(x = x, y= y), size = 2.5, alpha = 0.25, colour = "white") +
  geom_text(data = weights, aes(x = x, y= y, label = weight), size = 2.5, family = "Times") +
	scale_fill_manual(values = c("#636363", "white")) + 
	scale_colour_manual(values = c("#636363", "#bdbdbd")) +
	my_theme + xlab("") + ylab("") + 
  	xlim(c(-0.09, 2.09)) + ylim(c(-0.15, 1.15)) +
 theme(plot.margin=unit(c(4,4,-4,-4),"mm")) +
	scale_size_manual(values = c(0.5, 0.5)) +
  ggtitle("a) visitation frequency")

############################
sp_names <- c("a[1]", "a[2]", "a[3]", "p[1]", "p[2]")

adj_matrix <- c(0,0,0,1,0,
								0,0,0,1,1,
								0,0,0,1,1,
								1,1,1,0,0,
								0,1,1,0,0) %>%
	matrix(nrow = 5, byrow = T) %>%
	`rownames<-`(sp_names) %>%
	`colnames<-`(sp_names)

net_exp <- adj_matrix %>% 
	igraph::graph_from_adjacency_matrix(mode = "directed") 
igraph::V(net_exp)$type <- c(F, F, F, T, T)
igraph::E(net_exp)$weight <- 1

layoutCoordinates <- cbind(c(0,1,2,0.5,1.5), 
													 c(1,1,1,0,0)) %>%
	`colnames<-`(c("x", "y"))

adj_list <- adj_matrix %>% 
	reshape2::melt() %>% 
	dplyr::filter(value > 0)

# Generate an edge path for each pair of connected nodes
n <- 17
allEdges <- lapply(1:nrow(adj_list), edgeMaker, len = n, curved = FALSE)
allEdges <- do.call(rbind, allEdges) %>%
	dplyr::filter(Sequence != n) %>%
	dplyr::mutate(Group = factor(Group),
				 type = Group)

new_lev <- c("m", "u", "m", "u", "u", "u", "m", "u", "u", "m")
dir <- c("m", "m", "m", "m", "m", "u", "u", "u", "u", "u")
dir2 <- c("m", "u", "m", "u", "u", "u", "m", "m", "u", "m")
id <- 1:10
allEdges %<>% dplyr::mutate(dir = type,
                            dir2 = dir)
levels(allEdges$type) <- new_lev
levels(allEdges$dir) <- dir
levels(allEdges$dir2) <- dir2
allEdges %<>% 
  dplyr::mutate(dir = as.numeric(dir),
                dir2 = as.numeric(dir2),
                dir = plyr::mapvalues(dir, c(1,2), c(0.03, -0.03)),
                dir2 = plyr::mapvalues(dir2, c(1,2), c(0.03, -0.03)))

nodes <- data.frame(layoutCoordinates) %>%
	dplyr::mutate(spe = sp_names,
								spe_c = c("u", "m", "m", "m", "m"), 
								pos = rep(c(0.3, -0.3), c(3,2)))

# ggplot(allEdges) +  # Pretty simple plot code
# 	geom_path(aes(x = x + dir, y = y + dir2, group = Group, colour = type, size = type), 
# 						arrow = arrow(type = "closed", angle = 15, length = grid::unit(0.1, "inches")))

weights <- allEdges %>%
  dplyr::filter(Sequence == 12) %>%
  dplyr::mutate(weight = c("1", "0.5", "0.5", "0.5", "0.5", "0.25", "0.5", "0.25", "0.66", "0.33"))

p2 <- ggplot(allEdges) +  # Pretty simple plot code
	geom_path(aes(x = x + dir, y = y + dir2, group = Group, size = type), 
						arrow = arrow(type = "closed", angle = 15, length = grid::unit(0.1, "inches")), colour = "#bdbdbd") + # and taper
	geom_point(data = nodes,  # Add nodes
						 aes(x = x, y = y ), size = 5, pch = 21,
						 colour = "black", fill = "white") +  
	# geom_text(data = nodes, aes(x = x, y = y + pos, label = spe), size = 3, parse = T, family = "Times") +
  geom_point(data = weights, aes(x = x + dir*3, y= y), size = 2.5, alpha = 0.25, colour = "white") +
  geom_text(data = weights, aes(x = x + dir*3, y= y, label = weight), size = 2.5, family = "Times", parse = T) +
	scale_fill_manual(values = c("#636363", "white")) + 
	scale_colour_manual(values = c("#636363", "#bdbdbd")) +
	my_theme + xlab("") + ylab("") + 
  xlim(c(-0.09, 2.09)) + ylim(c(-0.15, 1.15)) +
 theme(plot.margin=unit(c(4,4,-4,-4),"mm")) +
	scale_size_manual(values = c(0.5, 0.5)) +
  ggtitle("b) mutual dependence")

####################################

sp_names <- c("a[1]", "a[2]", "a[3]", "p[1]", "p[2]")

adj_matrix <- c(0,0,0,0,0,
								0,0,0,1,1,
								0,0,0,0,0,
								1,1,1,0,0,
								0,0,1,0,0) %>%
	matrix(nrow = 5, byrow = T) %>%
	`rownames<-`(sp_names) %>%
	`colnames<-`(sp_names)

net <- adj_matrix %>% 
	igraph::graph_from_adjacency_matrix(mode = "directed") 
igraph::V(net)$type <- c(F, F, F, T, T)
igraph::E(net)$weight <- 1

layoutCoordinates <- cbind(c(0,1,2,0.5,1.5), 
													 c(1,1,1,0,0)) %>%
	`colnames<-`(c("x", "y"))

adj_list <- adj_matrix %>% 
	reshape2::melt() %>% 
	dplyr::filter(value > 0)

# Generate an edge path for each pair of connected nodes
n <- 17
allEdges <- lapply(1:nrow(adj_list), edgeMaker, len = n, curved = FALSE)
allEdges <- do.call(rbind, allEdges) %>%
	dplyr::filter(Sequence != n) %>%
	dplyr::mutate(Group = factor(Group),
				 type = Group)

# new_lev <- c("m", "u", "m", "m", "u")
# levels(allEdges$type) <- new_lev

nodes <- data.frame(layoutCoordinates) %>%
	dplyr::mutate(spe = sp_names,
								# spe_c = c("m", "m", "u", "m", "u"), 
								pos = rep(c(0.3, -0.3), c(3,2)))

weights <- allEdges %>%
  dplyr::filter(Sequence == 12) %>%
  dplyr::mutate(weight = c("0.75", " ", "0.5", "0.33", " ", "0.25"))

p3 <- ggplot(allEdges) +  # Pretty simple plot code
	geom_path(aes(x = x, y = y, group = Group), size = 0.5,
						arrow = arrow(type = "closed", angle = 15, length = grid::unit(0.1, "inches")), colour = "#bdbdbd") + # and taper
	geom_point(data = nodes,  # Add nodes
						 aes(x = x, y = y ), size = 5, pch = 21,
						 colour = "black", fill = "white") + 
	# geom_text(data = nodes, aes(x = x, y = y + pos, label = spe), size = 3, parse = T, family = "Times") +
   geom_point(data = weights, aes(x = x, y= y), size = 2.5, alpha = 0.25, colour = "white") +
  geom_text(data = weights, aes(x = x, y= y, label = weight), size = 2.5, family = "Times") +
	scale_fill_manual(values = c("#636363", "white")) + 
	scale_colour_manual(values = c("#636363", "#bdbdbd")) +
	my_theme + xlab("") + ylab("") + 
  	xlim(c(-0.09, 2.09)) + ylim(c(-0.15, 1.15)) +
 theme(plot.margin=unit(c(4,4,-4,-4),"mm")) +
	# scale_size_manual(values = c(0.5, 0.5)) +
  ggtitle("c) interaction asymmetry")

###########################

multiplot(p1, p2, p3, cols = 3)

@
\caption{\label{fig:explain_cd} Different ways to depict quantitative mutualistic networks. a) Pollination networks are usually described by the number of visits between each plant and animal species. b) Given visitation data the mutual dependencies between interacting species depend directly on the relative visitation frequency. c) The relative differences of dependencies---the interaction asymmetry---then provides a mean to simplify the interspecific effects.}
\end{figure}

\subsubsection*{Data analysis}

Although understanding the variability of manageability across ecological networks is an useful result itself, we also wanted to test whether network attributes, like invasion status, have an impact on the observed manageability values. We therefore used a set of generalised linear models (with binomial error structure) to investigate the effect of invasion status while controling for factors related to species richness, since one might naively expect to see a negative relationship between diversity and manageability \autocite{Menge1995}. These covariates inlcuded the total number of species, plant richness, polinators richness, the ratio of plants to pollinators richness, and the study site. We assessed competing models by comparing the Akaike Information Criterion corrected for small sample sizes (AICc), and report results for the most parsimonious one.

We next explored the extent to which structural properties of the empirical networks influence their manageability. Previous research indicates a direct link between a network's degree distribution and the number of nodes necessary to fully control it \autocite{Liu2011}, but the strength and applicability of this relationship has not been tested for in weighted ecological networks. We therefore compared the manageability of the empirical networks to networks generated by three null models which varied in terms of the constraints they impose. The first set of random visitation networks maintain the connectivity of the empirical network, but both degree (its number of interactions) and strength (its total sum of visits) were allowed to vary. A second set set was generated by maintaining each species' degree while varying the number of visits per species. Finally a third set of random networks mantained the species' strength while varying their degrees. Randomisations were generated using vegan 2.3-3 \autocite{Oksansen2016}. After generating the randomized networks under these three schemes, we then calculated the mutual dependences and interaction asymmetries of each and determined their manageability using the maximum matching algorithm. Finally we calculated the average rank (akin to a p-value) of the manageability of each empirical network compared to the corresponding manageabilites of each set of 999 randomisations.

>>>>

---
title: 
csl: ecology-letters.csl
output: html_document
bibliography: references.bib
---

**Figure 1:**
Finding a maximum matching in a complex network. 
a) We start with a directed network  in which the direction of the link represent the direction of control. 
This is, a link from $a_1$ to $p_1$ indicates that the state of $p_1$ (density/abundances in an ecological context) are influenced by the state of $a_1$. 
b) To facilitate the computation of the maximum matching, we can use an alternative representation of the directed network in which each node in (a) is represented by two nodes that indicate their outgoing and incoming links. 
In both (a) and (b) we show one of the possible maximum matchings of a network (dark grey links) that defines a set of matched ($p_1$, $p_2$, and $a_3$), unmatched ($a_1$ and $a_2$) and superior nodes ($a_1$, $a_2$, and $p_1$; dark grey circles). 
The minimum number of species neccessary to control this network equates to the number of unmatched nodes $n_d = 2$, which gives a manageability of $m = 0.8$.

**Figure 2:** 
Different ways to depict quantitative mutualistic networks. 
a) Pollination networks are usually described by the number of visits between each plant and animal species. 
b) Given visitation data the mutual dependencies between interacting species depend directly on the relative visitation frequency. 
c) The relative differences of dependencies---the interaction asymmetry---then provides a means to simplify the interspecific effects.

**Figure 3:**
Invaded communities have lower levels of manageability than uninvaded communities even when controling for the ratio of plant to pollinator richness. 
Adjusted manageability corrsponds to the partial working residuals of the invasion status. 
The box covers the 25th–75th percentiles, the middle line marks the median, and the maximum length of the whiskers is 1.5 times the interquartile range. 
Points outside this range show up as outliers.

**Figure 4:**
The manageability of empirical networks compared to network randomisations. 
For each of four randomisation approaches, we show that---akin to a p-value---an scaled mean rank $< 0.025$ or $> 0.975$ (shaded areas in light grey) suggest a significant difference between the empirical network and its randomisations. 
The empirical manageability is much smaller than those of randomisations in which the direction of assymmetries has been randomised. 
In contrast the manageability of network in which the visitation strength or degree was mantained was not significantly different to the manageability of the empirical networks. 
Dark grey and white boxplots represent manageabilities obtained using mutual dependencies and asymmetries respectively. 
All boxes are as in Figure 3.

**Figure 5:**
Dependency strength is the single most important factor explaining the importance of species for network control. 
Visitation levels and contribution to nestedness had a significant albeit small effect on the importance (dashed lines correspond to ± one standard deviation on these factors). 
Invasive species (red) were invariably classified as driver species (those having the maximum importance for network control.

**Table 1:**
The strength of dependency, contribution to nestedness and visitation strength had a significative effect on all the models that accounted for 95\% of the evidence based on AICc. .
